Introducing ICBe: An Event Extraction Dataset from Narratives about International Crises

### 2024-04-08 replication_paper/psrm_draft/ICBe_manuscript.Rmd
replication merge

#Execute these lines to get bbl files for the arxiv submission
options(tinytex.clean=F)
```

@@

```{=tex}
\clearpage
\singlespacing
```
```{r, eval=T, echo=F, results='asis', include=T, message=F, cache=F, warning=F, ft.arraystretch=1}

@@

library(googlesheets4) #Do this high up so you get the prompt early
library(lubridate, warn.conflicts = FALSE)
lit_review_pull <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/14RHRKgQGHj7Zr4_xt4fq2bIE7mDNayLKj89tYmOOfHQ/edit?usp=sharing", 
                                sheet="DatasetsComparisonLitReview") #%>% janitor::clean_names() 
# dim(lit_review_pull)

@@

@@

actor_translator <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1a7Id0Zg41PTKEv74H_KjiJzhMak2jGh0leT7B8oVWdI/edit?usp=sharing", 
                                sheet="actor_translator") #%>% janitor::clean_names() 
#Convert to a local file for pushing
#actor_translator    <- readxl::read_excel(target_file, sheet="actor_translator") #Ya we just add colors to the actor translator and we're good
pallet <- c(brewer.pal(n=20, "Dark2"), brewer.pal(n=20, "Set1"), brewer.pal(n=20, "Set2"), brewer.pal(n=20, "Paired")) %>% unique()
library(pals); #install.packages('pals')
pallet <- c(pals::glasbey(),  pals::alphabet2())
pallet <- setdiff(pallet, c("#666666", "#999999", "#FFFF99","#FFFF33","#B2DF8A","#B3B3B3","#0000FF")) 
library(colorspace)
#cols <- c("#CDE4F3","#E7F3D3","#F7F0C7","#EFCFE5","#D0D1E7")    
pallet <- rgb(hex2RGB(pallet)@coords * 0.7)
actor_translator$actor_color = rep(pallet, length=nrow(actor_translator))

@@

```{python sentence-transformers, eval=T, results='hide', echo=F, include=T, message=F, cache=F, warning=F}
#Can't get this and spacy to work in the same run. Once reticulate loads once it sticks with the environment that came first and I can't figure out how to switch it.
#pip install -U sentence-transformers

fromscratch=False

if fromscratch:
  from sentence_transformers import SentenceTransformer
  import numpy as np
  sentences=np.array(r.sentences_unique) #sentences_unique_sanitized
  sentence_chars = np.array([len(q) for q in sentences])
  #sentences=sentences[sentence_chars<1000] #cull anything with too many characters
  len(sentences) #12161 12162
  n=len(sentences)
  chunk_length=100
  n_chunks = int(np.ceil(n/chunk_length)) #can only 
  sentences_chunks=[sentences[i:i + chunk_length] for i in range(0, n, chunk_length)]
  len(sentences_chunks)
  len(sentences_chunks[0])  
  #model = SentenceTransformer('sentence-transformers/all-mpnet-base-v2')
  #model = SentenceTransformer('sentence-transformers/gtr-t5-xxl')
  #model = SentenceTransformer('sentence-transformers/gtr-t5-xl')
  model = SentenceTransformer('paraphrase-mpnet-base-v2')
  #sentences
  #sentences_test=['That is a happy person','That is a happy dog','That is a not happy person','That person is happy','This person is happy']
  #temp = model.encode(sentences_chunks[0])
  sentences_embeddings = np.vstack([model.encode(q) for q in sentences_chunks])  #[0:100] #cpu is possible just painfully slow #, device='0'
  sentences_embeddings.shape #(12103, 768)

else:
  print("Skipping")
```

```{r sentence-embeddings, results='hide', echo=F, include=T, message=F, cache=F, warning=F}
fromscratch = FALSE
if(fromscratch){
  sentences_embeddings <- py$sentences_embeddings
  dim(sentences_embeddings)
  sentences_embeddings %>% saveRDS(paste0(here::here(), '/replication_paper/data/temp/sentences_embeddings.Rds'))
}else{
  sentences_embeddings <- readRDS(paste0(here::here(), '/replication_paper/data/temp/sentences_embeddings.Rds'))
  dim(sentences_embeddings)
}
```

```{r umap, results='hide', echo=F, include=T, message=F, cache=F, warning=F, eval=F}
#The [umapr](https://github.com/ropensci-archive/umapr) package requires python and [umap](https://github.com/lmcinnes/umap#installing). Please check their repos for installation details.

@@

```{r sentence-embeddings-dendro, eval=F, results='hide', echo=F, include=T, message=F, cache=F, warning=F, ft.arraystretch=0.75}

@@

library(googlesheets4) #Do this high up so you get the prompt early
library(lubridate, warn.conflicts = FALSE)
case_study_cuban <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1NuRWFB1HEbQbJu_JoEzJ9WYk7kfwwCryF7wBfKPT6uc/edit?usp=sharing", 
                                sheet="final") %>% janitor::clean_names() 
dim(case_study_cuban)
#case_study_cuban$date <- case_study_cuban$date %>% lapply(FUN=function(x) { ifelse(is.null(x), "", x[[1]] %>% as.character() )   }   ) %>% unlist() %>% as.Date()

paste0(" ![](", here::here(), "/replication_data/in/flags_small//Q241.png) ")

temp1 <- case_study_cuban %>% 
                #filter(!is.na(date)) %>%
                #head(2) %>%
                mutate(date=date %>% anytime::anydate()) %>%
                mutate(year=year(date) %>% as.character() ) %>%
                mutate(month=month(date) %>% as.character()) %>%
                mutate(day=day(date) %>% as.character()) %>%
                mutate(documents = documents %>% as.numeric()) %>%
                mutate(Rank=rank(-documents, ties.method="first")  ) %>%
                dplyr::select(Y=year,M=month, D=day, Rank, Event=event, Docs=documents, icb_narrative=icb_narrative, ICBe=ic_be, ICB=icb, MIDs=mids, phoenix=phoenix) %>% 
                mutate(Event= Event %>%  replace_emoticons) %>% 
                mutate(icb_narrative = ifelse(!is.na(icb_narrative), "✓", NA)) %>%
                mutate(ICB = ifelse(!is.na(ICB) & !ICB %in% 0, "✓", NA)) %>%
                mutate(MIDs = ifelse(!is.na(MIDs) & !MIDs %in% 0, "✓", NA)) %>%
                mutate(phoenix = ifelse(!is.na(phoenix) & !phoenix %in% 0, "✓", NA)) %>%
                mutate(ICBe = ifelse(!is.na(ICBe) & !ICBe %in% 0, "✓", NA)) %>%

#Ok rather than try to match up we're going to do this by hand now

```{r recall-crimea, eval=T, results='hide', echo=F, include=T, message=F, cache=F, warning=F, ft.arraystretch=0.75}
library(dplyr)
library(googlesheets4) #Do this high up so you get the prompt early
library(lubridate, warn.conflicts = FALSE)
case_study_crimea <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1YesAx1CkYCgrEi_WVJ9aho60HesWA6p-3XMma9IzBrI/edit?usp=sharing", 
                                sheet="final") %>% janitor::clean_names() 
dim(case_study_crimea)

rex_date <- function(x){
  a <- x[1] %>% as.character() %>% unlist()
  if(is.null(a)){return('')}
  return(a)
}
case_study_crimea$date <- sapply(sapply(case_study_crimea$date, as.character), rex_date)
temp2 <- case_study_crimea %>% 
                filter(keep %in% 1) %>%
                #filter(!is.na(date)) %>%
                #head(2) %>%
                mutate(date=date %>% anytime::anydate()) %>%
                mutate(year=year(date) %>% as.character() ) %>%
                mutate(month=month(date) %>% as.character()) %>%
                mutate(day=day(date) %>% as.character()) %>%
                mutate(n_sources = n_sources %>% as.numeric()) %>%
                mutate(Rank=rank(-n_sources, ties.method="first")  ) %>%
                dplyr::select(Y=year,M=month, D=day, Rank,  Event=event, Docs=n_sources, 
                              icb_narrative=icb_narrative, ICBe=ic_be, MIDs=mids, phoenix, terrier, icews ) %>%  #, ICB=icb, MIDs=mids, phoenix=phoenix
                mutate(Event= Event %>%  replace_emoticons_crimea) %>%
                arrange(Y %>% as.numeric() ,M %>% as.numeric()  ,D %>% as.numeric() ) %>%
                mutate(D=ifelse(!is.na(lag(D)) & D==lag(D) & M==lag(M),'',D )) %>%
                mutate(M=ifelse(!is.na(lag(M)) & M==lag(M),'',M )) %>%
                mutate(Y=ifelse(!is.na(lag(Y)) & Y==lag(Y),'',Y )) %>%
                mutate(Y=ifelse(duplicated(Y), NA, Y)) %>%
                mutate(YMD=paste(Y,
                                 M %>% str_pad(2, pad = "0") ,
                                 D %>% str_pad(2, pad = "0"), 
                                 sep="-") %>% trimws() %>%
                                 str_replace_all("^NA-","") %>%  #There's like a carriage return and tab in there somehow!
                                 str_replace_all("^NA-","") %>% 
                                 str_replace_all("^-00|^00","") %>% 
                                 str_replace_all("^-00|^00","") %>% 
                                 str_replace_all("^-","") 
                       ) %>%
                dplyr::select(-Y,-M,-D) %>%
                relocate(YMD, .before = Rank) %>%
                dplyr::select(-Rank) %>% 
                mutate(icb_narrative = ifelse(!is.na(icb_narrative), "✓", NA)) %>%
                mutate(icews = ifelse(!is.na(icews) & !icews %in% 0 , "✓", NA)) %>%
                mutate(MIDs = ifelse(!is.na(MIDs) & !MIDs %in% 0  , "✓", NA)) %>%
                mutate(phoenix = ifelse(!is.na(phoenix) & !phoenix %in% 0  , "✓", NA)) %>%
                mutate(terrier = ifelse(!is.na(terrier) & !terrier %in% 0  , "✓", NA))  %>%
                mutate(ICBe = ifelse(!is.na(ICBe) & !ICBe %in% 0  , "✓", NA))

@@

```{r precision, eval=F, results='hide', echo=F, include=T, message=F, cache=F, warning=F, ft.arraystretch=0.75}

>>>>

#Execute these lines to get bbl files for the arxiv submission
options(tinytex.clean=F)
fromscratch = FALSE #Set to TRUE to repull google docs
```

@@

```{=tex}
\clearpage
\singlespacing
```

```{r, eval=T, echo=F, results='asis', include=T, message=F, cache=F, warning=F, ft.arraystretch=1}

@@

library(googlesheets4) #Do this high up so you get the prompt early
library(lubridate, warn.conflicts = FALSE)

if(fromscratch){googlesheets4::gs4_deauth()}

if(fromscratch){
  lit_review_pull <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/14RHRKgQGHj7Zr4_xt4fq2bIE7mDNayLKj89tYmOOfHQ/edit?usp=sharing", sheet="DatasetsComparisonLitReview") #%>% janitor::clean_names() 
  saveRDS(lit_review_pull ,
            file=paste0(here::here(),"/replication_data/out/lit_review_pull.Rds")) 
} else {lit_review_pull <- readRDS( paste0(here::here(), "/replication_data/out/lit_review_pull.Rds"))}

# dim(lit_review_pull)

@@

library(googlesheets4) #Do this high up so you get the prompt early
library(lubridate, warn.conflicts = FALSE)

if(fromscratch){
  case_study_cuban <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1NuRWFB1HEbQbJu_JoEzJ9WYk7kfwwCryF7wBfKPT6uc/edit?usp=sharing", sheet="final") %>% janitor::clean_names() 
  saveRDS(case_study_cuban, file=paste0(here::here(),"/replication_data/out/case_study_cuban.Rds")) 
} else {case_study_cuban <- readRDS( paste0(here::here(), "/replication_data/out/case_study_cuban.Rds"))}

dim(case_study_cuban)
#case_study_cuban$date <- case_study_cuban$date %>% lapply(FUN=function(x) { ifelse(is.null(x), "", x[[1]] %>% as.character() )   }   ) %>% unlist() %>% as.Date()

temp1 <- case_study_cuban %>% 
                #filter(!is.na(date)) %>%
                #head(2) %>%
                mutate(date=date %>% anytime::anydate()) %>%
                mutate(year=year(date) %>% as.character() ) %>%
                mutate(month=month(date) %>% as.character()) %>%
                mutate(day=day(date) %>% as.character()) %>%
                mutate(documents = documents %>% as.numeric()) %>%
                mutate(Rank=rank(-documents, ties.method="first")  ) %>%
                dplyr::select(Y=year,M=month, D=day, Rank, Event=event, Docs=documents, icb_narrative=icb_narrative, ICBe=ic_be, ICB=icb, MIDs=mids, phoenix=phoenix) %>% 
                # mutate(Event= Event %>%  replace_emoticons) %>% 
                mutate(icb_narrative = ifelse(!is.na(icb_narrative), "✓", NA)) %>%
                mutate(ICB = ifelse(!is.na(ICB) & !ICB %in% 0, "✓", NA)) %>%
                mutate(MIDs = ifelse(!is.na(MIDs) & !MIDs %in% 0, "✓", NA)) %>%
                mutate(phoenix = ifelse(!is.na(phoenix) & !phoenix %in% 0, "✓", NA)) %>%
                mutate(ICBe = ifelse(!is.na(ICBe) & !ICBe %in% 0, "✓", NA)) %>%

```{r recall-crimea, eval=T, results='hide', echo=F, include=T, message=F, cache=F, warning=F, ft.arraystretch=0.75}
library(dplyr)
library(googlesheets4) #Do this high up so you get the prompt early
library(lubridate, warn.conflicts = FALSE)

if(fromscratch){
  case_study_crimea <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1YesAx1CkYCgrEi_WVJ9aho60HesWA6p-3XMma9IzBrI/edit?usp=sharing", sheet="final") %>% janitor::clean_names() 
  saveRDS(case_study_crimea, file=paste0(here::here(),"/replication_data/out/case_study_crimea.Rds")) 
} else {case_study_crimea <- readRDS( paste0(here::here(), "/replication_data/out/case_study_crimea.Rds"))}

dim(case_study_crimea)

rex_date <- function(x){
  a <- x[1] %>% as.character() %>% unlist()
  if(is.null(a)){return('')}
  return(a)
}
case_study_crimea$date <- sapply(sapply(case_study_crimea$date, as.character), rex_date)
temp2 <- case_study_crimea %>% 
                filter(keep %in% 1) %>%
                #filter(!is.na(date)) %>%
                #head(2) %>%
                mutate(date=date %>% anytime::anydate()) %>%
                mutate(year=year(date) %>% as.character() ) %>%
                mutate(month=month(date) %>% as.character()) %>%
                mutate(day=day(date) %>% as.character()) %>%
                mutate(n_sources = n_sources %>% as.numeric()) %>%
                mutate(Rank=rank(-n_sources, ties.method="first")  ) %>%
                dplyr::select(Y=year,M=month, D=day, Rank,  Event=event, Docs=n_sources, 
                              icb_narrative=icb_narrative, ICBe=ic_be, MIDs=mids, phoenix, terrier, icews ) %>%  #, ICB=icb, MIDs=mids, phoenix=phoenix
                # mutate(Event= Event %>%  replace_emoticons_crimea) %>%
                arrange(Y %>% as.numeric() ,M %>% as.numeric()  ,D %>% as.numeric() ) %>%
                mutate(D=ifelse(!is.na(lag(D)) & D==lag(D) & M==lag(M),'',D )) %>%
                mutate(M=ifelse(!is.na(lag(M)) & M==lag(M),'',M )) %>%
                mutate(Y=ifelse(!is.na(lag(Y)) & Y==lag(Y),'',Y )) %>%
                mutate(Y=ifelse(duplicated(Y), NA, Y)) %>%
                mutate(YMD=paste(Y,
                                 M %>% str_pad(2, pad = "0") ,
                                 D %>% str_pad(2, pad = "0"), 
                                 sep="-") %>% trimws() %>%
                                 str_replace_all("^NA-","") %>%  #There's like a carriage return and tab in there somehow!
                                 str_replace_all("^NA-","") %>% 
                                 str_replace_all("^-00|^00","") %>% 
                                 str_replace_all("^-00|^00","") %>% 
                                 str_replace_all("^-","") 
                       ) %>%
                dplyr::select(-Y,-M,-D) %>%
                relocate(YMD, .before = Rank) %>%
                dplyr::select(-Rank) %>% 
                mutate(icb_narrative = ifelse(!is.na(icb_narrative), "✓", NA)) %>%
                mutate(icews = ifelse(!is.na(icews) & !icews %in% 0 , "✓", NA)) %>%
                mutate(MIDs = ifelse(!is.na(MIDs) & !MIDs %in% 0  , "✓", NA)) %>%
                mutate(phoenix = ifelse(!is.na(phoenix) & !phoenix %in% 0  , "✓", NA)) %>%
                mutate(terrier = ifelse(!is.na(terrier) & !terrier %in% 0  , "✓", NA))  %>%
                mutate(ICBe = ifelse(!is.na(ICBe) & !ICBe %in% 0  , "✓", NA))

@@

if(fromscratch){
  actor_translator <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1a7Id0Zg41PTKEv74H_KjiJzhMak2jGh0leT7B8oVWdI/edit?usp=sharing", sheet="actor_translator") #%>% janitor::clean_names() 
  saveRDS(actor_translator,
            file=paste0(here::here(),"/replication_data/out/actor_translator.Rds")) 
} else {actor_translator <- readRDS( paste0(here::here(), "/replication_data/out/actor_translator.Rds"))}

#Convert to a local file for pushing
#actor_translator    <- readxl::read_excel(target_file, sheet="actor_translator") #Ya we just add colors to the actor translator and we're good
pallet <- c(brewer.pal(n=20, "Dark2"), brewer.pal(n=20, "Set1"), brewer.pal(n=20, "Set2"), brewer.pal(n=20, "Paired")) %>% unique()
library(pals); #install.packages('pals')
pallet <- c(pals::glasbey(),  pals::alphabet2())
pallet <- setdiff(pallet, c("#666666", "#999999", "#FFFF99","#FFFF33","#B2DF8A","#B3B3B3","#0000FF")) 
library(colorspace)
#cols <- c("#CDE4F3","#E7F3D3","#F7F0C7","#EFCFE5","#D0D1E7")    
pallet <- rgb(hex2RGB(pallet)@coords * 0.7)
actor_translator$actor_color = rep(pallet, length=nrow(actor_translator))

@@

```{python sentence-transformers, eval=fromscratch, results='hide', echo=F, include=T, message=F, cache=F, warning=F}
#Can't get this and spacy to work in the same run. Once reticulate loads once it sticks with the environment that came first and I can't figure out how to switch it.
#pip install -U sentence-transformers
  from sentence_transformers import SentenceTransformer
  import numpy as np
  sentences=np.array(r.sentences_unique) #sentences_unique_sanitized
  sentence_chars = np.array([len(q) for q in sentences])
  #sentences=sentences[sentence_chars<1000] #cull anything with too many characters
  len(sentences) #12161 12162
  n=len(sentences)
  chunk_length=100
  n_chunks = int(np.ceil(n/chunk_length)) #can only 
  sentences_chunks=[sentences[i:i + chunk_length] for i in range(0, n, chunk_length)]
  len(sentences_chunks)
  len(sentences_chunks[0])  
  #model = SentenceTransformer('sentence-transformers/all-mpnet-base-v2')
  #model = SentenceTransformer('sentence-transformers/gtr-t5-xxl')
  #model = SentenceTransformer('sentence-transformers/gtr-t5-xl')
  model = SentenceTransformer('paraphrase-mpnet-base-v2')
  #sentences
  #sentences_test=['That is a happy person','That is a happy dog','That is a not happy person','That person is happy','This person is happy']
  #temp = model.encode(sentences_chunks[0])
  sentences_embeddings = np.vstack([model.encode(q) for q in sentences_chunks])  #[0:100] #cpu is possible just painfully slow #, device='0'
  sentences_embeddings.shape #(12103, 768)
```

```{r sentence-embeddings, results='hide', echo=F, include=T, message=F, cache=F, warning=F}
if(fromscratch){
  sentences_embeddings <- py$sentences_embeddings
  dim(sentences_embeddings)
  sentences_embeddings %>% saveRDS(paste0(here::here(), '/replication_paper/data/temp/sentences_embeddings.Rds'))
}else{
  sentences_embeddings <- readRDS(paste0(here::here(), '/replication_paper/data/temp/sentences_embeddings.Rds'))
  dim(sentences_embeddings)
}
```

```{r umap, results='hide', echo=F, include=T, message=F, cache=F, warning=F, eval=fromscratch}
#The [umapr](https://github.com/ropensci-archive/umapr) package requires python and [umap](https://github.com/lmcinnes/umap#installing). Please check their repos for installation details.

@@

```{r sentence-embeddings-dendro, eval=fromscratch, results='hide', echo=F, include=T, message=F, cache=F, warning=F, ft.arraystretch=0.75}

@@

@@

```{r precision, eval=fromscratch, results='hide', echo=F, include=T, message=F, cache=F, warning=F, ft.arraystretch=0.75}
